---
title: "Introduction to Two Functions"
author: "Hao Yin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Two Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

__Hybrid.PSO__ is a function to implement Hybrid Particle Swarm Optimization Algorithm using R.

__PageRank__ is a function to implement PageRank Algorithm using Rcpp.

---

## Hybrid Particle Swarm Optimization Algorithm

### Algorithm Background

particle swarm optimization (PSO) is a computational method that optimizes a problem by iteratively trying to improve a candidate solution with regard to a given measure of quality. 

It solves a problem by having a population of candidate solutions, here dubbed particles, and moving these particles around in the search-space according to simple mathematical formula over the particle's position and velocity. 

Each particle's movement is influenced by its local best known position, but is also guided toward the best known positions in the search-space, which are updated as better positions are found by other particles. This is expected to move the swarm toward the best solutions.


### Source Code for the algorithm

The source R code for __Hybrid.PSO__ is as follows: 

```{r,eval=FALSE}
Hybrid.PSO <- function(solveFun,particleNum,limitX,w,c1,c2,iters,alpha){
  pbest <- NULL
  gbest <- NULL
  gbestAdd <- NULL
  vmax <- 0.15 * (limitX[2] - limitX[1])
  xMat <- matrix(c(x = runif(particleNum, limitX[1], limitX[2])), dimnames = list(NULL, c("x")))
  vMat <- matrix(c(x = runif(particleNum, -vmax, vmax)),  dimnames = list(NULL, c("x")))
  adjusts <- apply(xMat, 1, solveFun)
  pbest <- cbind(xMat, adjusts)
  idxAdjusts <- ncol(pbest)
  gbest <- pbest[which.max(pbest[, idxAdjusts]),]
  for (i in 1:iters){
    mapply(function(no, adj){
      if(adj > pbest[no, idxAdjusts]){
        pbest[no, ] <<- c(xMat[no, ], adj)
      }
    }, 1:length(adjusts), adjusts)
    
    if (max(pbest[, idxAdjusts]) > gbest[idxAdjusts]) {
      gbestAdd <- max(pbest[, idxAdjusts]) - gbest[idxAdjusts]
      gbest <- pbest[which.max(pbest[, idxAdjusts]), ]
    }
    xMatOld <- xMat
    xMat <- xMat + vMat
    vMat <- w*vMat +  
      c1 * runif(1, 0, 1) * (pbest[, 1:(idxAdjusts - 1), drop=F] - xMatOld) + 
      c2 * runif(1, 0, 1) * (matrix(rep(gbest[1:(idxAdjusts - 1)], particleNum), ncol = idxAdjusts - 1 , byrow = T)-xMatOld) 
    vMat[which(vMat < (-vmax))] <- (-vmax)
    vMat[which(vMat > (vmax))]  <- (vmax)
    xMat[which(xMat < (limitX[1]))] <- (limitX[1])
    xMat[which(xMat > (limitX[2]))] <- (limitX[2])
    adjusts <- apply(xMat, 1, solveFun)
    if (!is.null(gbestAdd) && gbestAdd < alpha) {
      return(gbest)
    }
  }
  cat("The iteration limit was reached and the result could not be solved")
}

```

### Algorithm application example

Here, we use a hybrid particle swarm optimization algorithm to solve the target function as following.
$$
f(x) = x*(sin(10*\pi*x)+2).
$$
We can see from the result graph that the HPSO Algorithm successfully find the optimal solution. 

```{r, fig.align='center'}
solveFun <- function(x){
  x*(sin(10*pi*x)+2)
}
limitX <- c(-1, 2) 
particleNum <- 20 
w <- 1          
c1 <- c2 <- 2   
iters <- 10000  
alpha <- 0.0002
solution <- Hybrid.PSO(solveFun,particleNum,limitX,w,c1,c2,iters,alpha)
cat("The optimal solution of the target function is x=", solution[1], ", f(x)=",solution[2], ".")
x <- seq(-1, 2, 0.01)
plot(x, solveFun(x), type = "l")
grid(col='grey60')
points(solution[1],solution[2],col="red")
```

---

## PageRank Algorithm

### Algorithm Background

PageRank (PR) is an algorithm used by Google Search to rank web pages in their search engine results. It is named after both the term "web page" and co-founder Larry Page. PageRank is a way of measuring the importance of website pages. 

PageRank works by counting the number and quality of links to a page to determine a rough estimate of how important the website is. The underlying assumption is that more important websites are likely to receive more links from other websites.

### Source Code for the algorithm

The source R code for __PageRank__ is as follows: 

```{r eval=FALSE}
NumericVector PageRank(NumericMatrix MA, double d, int MaxIter, NumericVector IniVec){
  
  int N = MA.nrow();
  NumericMatrix MT(N, N);
  NumericVector UpdVec(N);
  
  for(int i = 0; i < N; i++){
    double sum = 0;
    for(int j = 0; j < N; j++){
      sum = sum + MA(j,i);
    }
    if(sum>0){
      for(int k = 0; k < N; k++){
        MT(k,i) = MA(k,i)/sum;
      }
    }
  }
  
  for(int i = 0; i < MaxIter; i++){
    for(int j = 0; j < N; j++){
      UpdVec[j] = 0;
      for(int k = 0; k < N; k++){
        UpdVec[j] = UpdVec[j] + d*MT(j,k)*IniVec[k];
      }
      UpdVec[j] = UpdVec[j] + (1-d)/N;
    }
    double Diff = 0;
    for(int l = 0; l < N; l++){
      Diff = Diff + abs(UpdVec[l]-IniVec[l]);
    }
    if(Diff < 1e-10){
      break;
    }
    for(int m = 0; m < N; m++){
      IniVec[m] = UpdVec[m];
    }
  }
  
  return(UpdVec);
}
```

### Algorithm application example

We use a simulated adjacent matrix to illustrate the application result. Data 'MA' is a 50*50 adjacent matrix of 50 nodes. The PageRank algorithm give the importance evaluation results of each node. A bar figure is used to show the importance of each node.  

```{r fig.align='center'}
library(Rcpp)
library(ggplot2)
library(StatComp22035)

data(MA)
NumNode <- 50
nrow <- ncol <- NumNode
IniVec <- runif(NumNode, 0, 1)
Score <- PageRank(MA, 0.85, 100000, IniVec)
data <- data.frame(
  index = 1:NumNode,  
  value = Score
)
ggplot(data, aes(x=index, y=value)) + 
  geom_bar(stat = "identity", width=0.5)

```